---
title: "DS311 - R Lab Assignment"
author: "Maximilian Leitschuh"
date: "8/22/2022"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
variables <- ncol(mtcars)
observations <- nrow(mtcars) 
# Answer:
print(paste("There are total of", variables, "variables and", observations,"observations in this data set."))


#b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summ <- summary(mtcars)
summ

# Answer:
print("There are 5 discrete variables and 6 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- round(mean(mtcars$mpg), digits=2)
v <- round(var(mtcars$mpg), digits=2)
s <- round(sd(mtcars$mpg), digits=2)

print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ", v , " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
tab1 <- aggregate(mtcars$mpg, list(cyl=mtcars$cyl), mean)
tab1

tab2 <- aggregate(mtcars$mpg, list(gear=mtcars$gear), sd)
tab2
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
cross <- crosstable(mtcars, cols=c(cyl), by=c(gear), total = "row")
cross
print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
ggplot(PlantGrowth, aes(x=reorder(group,-weight), y=weight), fun.y = "mean") +
  geom_bar(stat="identity", fill="steelblue")+
  labs(y="mean weight", x="groups")+
  ggtitle("Mean weight of the groups")+
  scale_y_continuous(breaks = seq(0,50,10))+
  theme(plot.title = element_text(hjust=0.5, face="bold"))

```

Result:

=> On average, plants in group "trt2" weigh the most at just under 55 pounds. This is followed by the plants of group "ctrl" with about 50 pounds and the lightest are the plants of group "trt1" with an average of 46 pounds. 



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg,
     col='steelblue',
     main='Histogram of mpg',
     xlab='mpg',
     breaks=10,
     ylab='Frequency')


print("Most of the cars in this data set are in the class of 15 miles per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(~ Murder + Assault, data = USArrests, main = "Correlation between Murder and Assault")   

```

Result:

=> The correlations within the pairplot indicate that the more people arrested for assault, the more people arrested for murder, and vice versa. In my opinion, this is logical, because with an increase in arrests for assault, people's propensity to violence continues to increase.


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='housing.csv', mode='wb')
housingData <- read.csv('housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!


```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!


```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!


